Pattern matching for $(231,213)$-avoiding permutations

### 2015-01-30 article.tex
correct biblio

\tableofcontents

@@

	We consider here the so-called \emph{pattern containment} problem
	(also sometimes referred to as the pattern involvement problem):
	Given two permutations $\sigma$ and $\pi$, this problem is to decide whether
	$\sigma \preceq \pi$ (the problem is ascribed to Wilf in \cite{Bose:Buss:Lubiw:1998}).
	The permutation containment problem is \NPhard~\cite{Bose:Buss:Lubiw:1998}.
	It is, however, polynomial time solvable by brute-force enumeration
	if $\sigma$ has bounded size.
	Improvements to this algorithm were presented in
	\cite{Albert:Aldred:Atkinson:Holton:ISAAC:2001} and
	\cite{Ahal:Rabinovich:2008},
	the latter describing a $O(|\pi|^{0.47k+o(|\sigma| )})$ time algorithm.
	Bruner and Lackner \cite{Bruner:Lackner:SWAT:2012}
	gave a fixed-parameter algorithm solving the pattern containment problem with
	an exponential worst-case runtime of $O(1.79^{\RUN(\pi)})$,
	where $\RUN(\pi)$ denotes the number of alternating runs of $\pi$.
	(This is an improvement upon the $O(2^{|\pi|})$ runtime required by
	brute-force search without imposing restrictions on $\sigma$ and $\pi$.)

@@

	Of particular interest in our context,
	the pattern containment problem is also solvable in
	polynomial time for separable patterns.
	Separable permutations are those permutations that contain neither
	$2413$ nor $3142$, and they are enumerated by the Schröder numbers
	(sequence A006318 in OEIS).
	Separable permutations has gained a lot of interest in the pattern containment
	context during the last years.
	Indeed, the pattern containment problem is solvable $O(kn^4)$ time and
	$O(kn^3)$ space for separable patterns \cite{Ibarra:1997}
	(see also \cite{Bose:Buss:Lubiw:1998}),
	where $k$ is the length of the pattern and $n$ is the length of the
	target permutation.
	Notice that there are numerous characterizations of separable permutations.
	To mention just a few examples,
	they are the permutations whose permutation graphs are cographs
	(\emph{i.e.} $P_4$-free graphs);
	equivalently, a separable permutation is a permutation that can be obtained
	from the trivial permutation $1$ by \emph{direct sums} and \emph{skew sums}.
	To the best of our knowledge,
	separable permutations first arose in the work of
	Avis and Newborn~\cite{Avis:Monroe:1981},
	who showed that they are precisely the permutations which can be sorted by an
	arbitrary number of pop-stacks in series,
	where a pop-stack is a restricted form of stack in which any pop operation
	pops all items at once.
	(Notice that the separable permutations include as a special case the
	stack-sortable permutations, which avoid the pattern $231$.)
	Separable permutations
	have been widely studied in the last decade, both from a combinatorial
	\cite{West:1995} and an algorithmic
	\cite{Ibarra:1997,Bose:Buss:Lubiw:1998,Rossin:Bouvel:06,Bouvel:Rossin:Vialette:CPM:2007} point
	of view.

>>>>

@@

	We consider here the so-called \emph{pattern containment} problem
	(also sometimes referred to as the \emph{pattern involvement problem}):
	Given two permutations $\sigma$ and $\pi$, this problem is to decide whether
	$\sigma \preceq \pi$ (the problem is ascribed to Wilf in \cite{Bose:Buss:Lubiw:1998}).
	The permutation containment problem is \NPhard~\cite{Bose:Buss:Lubiw:1998}.
	It is, however, polynomial-time solvable by brute-force enumeration
	if $\sigma$ has bounded size.
	Improvements to this algorithm were presented in
	\cite{Albert:Aldred:Atkinson:Holton:ISAAC:2001} and
	\cite{Ahal:Rabinovich:2008},
	the latter describing a $O(|\pi|^{0.47k+o(|\sigma| )})$ time algorithm.
	Bruner and Lackner \cite{Bruner:Lackner:SWAT:2012}
	gave a fixed-parameter algorithm solving the pattern containment problem with
	an exponential worst-case runtime of $O(1.79^{\RUN(\pi)})$,
	where $\RUN(\pi)$ denotes the number of alternating runs of $\pi$.
	(This is an improvement upon the $O(2^{|\pi|})$ runtime required by
	brute-force search without imposing restrictions on $\sigma$ and $\pi$.)
	Of particular importance, it was recently proved that
	the permutation containment problem is fiexed-parameter tractable for
	parameter $|\sigma|$
	\cite{Guillemot:Marx:SODA:2014}.

@@

	Of particular interest in our context,
	the pattern containment problem is also solvable in
	polynomial time for separable patterns.
	Separable permutations are those permutations that contain neither
	$2413$ nor $3142$, and they are enumerated by the Schröder numbers
	(sequence A006318 in OEIS).
	Separable permutations has gained a lot of interest in the pattern containment
	context during the last years.
	Indeed, the pattern containment problem is solvable $O(kn^4)$ time and
	$O(kn^3)$ space for separable patterns \cite{Ibarra:1997}
	(see also \cite{Bose:Buss:Lubiw:1998}),
	where $k$ is the length of the pattern and $n$ is the length of the
	target permutation.
	Notice that there are numerous characterizations of separable permutations.
	To mention just a few examples,
	they are the permutations whose permutation graphs are cographs
	(\emph{i.e.} $P_4$-free graphs);
	equivalently, a separable permutation is a permutation that can be obtained
	from the trivial permutation $1$ by \emph{direct sums} and \emph{skew sums}.
	To the best of our knowledge,
	separable permutations first arose in the work of
	Avis and Newborn~\cite{Avis:Newborn:1981},
	who showed that they are precisely the permutations which can be sorted by an
	arbitrary number of pop-stacks in series,
	where a pop-stack is a restricted form of stack in which any pop operation
	pops all items at once.
	(Notice that the separable permutations include as a special case the
	stack-sortable permutations, which avoid the pattern $231$.)
	Separable permutations
	have been widely studied in the last decade, both from a combinatorial
	\cite{West:1995} and an algorithmic
	\cite{Ibarra:1997,Bose:Buss:Lubiw:1998,Rossin:Bouvel:06,Bouvel:Rossin:Vialette:CPM:2007} point
	of view.
